% !TEX root =  ../Report.tex
\section{Literature Review}
\label{sec: Literature Review}
% =============== BITCOIN AS FINANCIAL TIME SERIES ===============
\subsection{Bitcoin as a Financial Time Series}
Bitcoin’s statistical properties distinguish it from conventional financial assets.Tang et al. (2025) confirm through systematic high-frequency analysis that Bitcoin displays non-linear behaviour, characterised by non-gaussian errors, time-varying volatility, and short-range autocorrelation in squared returns. These properties are consistent with those of mature equity markets, but amplified in magnitude \cite{tangStylizedFactsHighFrequency2025}. Bouri et al. (2022) deduce that Bitcoin's maximum conditional volatility reaches two orders of magnitude above the S\&P 500 equivalent, with the tail index varying between approximately 2 and 3 over time \cite{bouriBitcoinSP500Comovements2022}. Empirical analysis of daily returns conducted by the Nakamoto Portfolio (2023) provides further evidence, reporting a Bitcoin kurtosis of 105 against the S\&P 500's approximately 10 \cite{BitcoinHistoricalReturns}. Critically, these studies demonstrate that Bitcoin's return distribution is non-Gaussian, making any model that assumes normality structurally misspecified.\\
\noindent The weak-form Efficient Market Hypothesis (EMH) implies that historical price series contain no exploitable predictive signal - directly challenging the premise of any ML-based trading model. Demmler and Lengyel-Almost (2021) found that 80\% of the 25 reviewed articles reject weak-form EMH for Bitcoin. Critically, the paper notes that the three studies accepting efficiency used data prior to the 2017 price surge \cite{lengyel-almosBitcoinMarketEfficient2021}. This raises the concern that their sample periods did not include the most extreme inefficiency episodes, potentially skewing their reasoning. Yi et al. (2023) quantified this nuance using daily data from 2010 to 2019. First, Bitcoin's excess kurtosis for annualised log returns was $\approx9$ times greater than S\&P 500. This, combined with the Jarque-Bera test on sampling intervals of 1 to 16 days rejecting normality at the 0.1\% significance level, is consistent with the distributional findings discussed above. The random walk probabilites in Table \ref{tab:lit_bitcoin_facts} demonstrate that Bitcoin's efficiency is time-varying, with speculative regime transitions characterised by underlying trends more than stable periods \cite{yiMarketEfficiencyCryptocurrency2023}.
\
{  
% \setlength{\LTleft}{-5pt}
\setlength{\LTright}{\fill}
\setlength{\LTpre}{-1pt}
\setlength{\LTpost}{-5pt}
  
\renewcommand{\arraystretch}{0.75}
\hyphenpenalty=50
\begin{longtable}{| 
    >{\raggedright\arraybackslash\leftskip=-4.5pt}p{0.19\textwidth} | 
    >{\raggedright\arraybackslash\leftskip=-4.5pt}p{0.37\textwidth} | 
    >{\raggedright\arraybackslash\leftskip=-4.5pt}p{0.35\textwidth} |}
\caption{Literature findings on Bitcoin stylised facts.\label{tab:lit_bitcoin_facts}}\\
\hline
\textbf{Citation} & 
\textbf{Methodology Used} & 
\textbf{Key Finding} \\ 
\hline
\endfirsthead
    
\hline
\multicolumn{3}{r}{} \\
\endfoot
    
\hline
\endlastfoot

Nakamoto Portfolio (2023) \cite{BitcoinHistoricalReturns} & 
Descriptive statistical analysis of daily Bitcoin and S\&P 500 returns & 
Bitcoin skewness = +2.8, kurtosis $\approx 105$; S\&P 500 skewness $\approx -0.3$, kurtosis $\approx 10$ \\ 
\hline
Bouri et al. (2022) \cite{bouriBitcoinSP500Comovements2022} & 
Conditional moments framework to daily Bitcoin and S\&P 500 returns; time-varying tail index estimation; higher-moment co-movement analysis & 
Bitcoin conditional volatility maximum is two orders of magnitude higher than S\&P 500; conditional skewness experiences far larger swings; tail index varies over time from $\mu \approx 2$ to $\mu \approx 3$ \\ 
\hline
Grobys (2023) \cite{grobysFractalComparativeView2023} & 
Estimate Hurst exponents for Bitcoin and S\&P 500 using Mandelbrot's Rescaled Range (R/S) analysis; Detrended Fluctuation Analysis (DFA) as robustness check & 
Bitcoin exhibits higher long-term dependence than S\&P 500; finite-kurtosis hypothesis rejected for S\&P 500 returns \\ 
\hline
\mbox{Demmler \& Lengyel}-Almos (2021) \cite{lengyel-almosBitcoinMarketEfficient2021} & 
Review and tabulation of EMH test conclusions from 25 peer-reviewed Bitcoin studies & 
80\% rejected weak-form EMH for Bitcoin \\ 
\hline
Yi et al. (2023) \cite{yiMarketEfficiencyCryptocurrency2023} & 
Quantum harmonic oscillator (QHO) efficiency framework - daily Bitcoin, gold, S\&P 500, and USD/EUR data (2010 to 2019); Jarque-Bera normality tests (1--16 days intervals); variance ratio tests & 
Excess kurtosis: BTC-42.5, S\&P 500-4.9, Gold-7.2, USD/EUR-1.8; Bitcoin random-walk P: 90.7\% overall, down to 87--88\% in 2013/2017 bubble episodes \\ 
\hline
\end{longtable}
}
\renewcommand{\arraystretch}{1.0}
% =============== METHODS FOR FINANCIAL TIME SERIES ANALYSIS ===============
\subsection{Methods for Financial Time Series Analysis}
\textbf{Autoregressive Integrated Moving Average (ARIMA):} ARIMA models the conditional mean as a linear function of past observations \cite{box1970time}. Various sources report ARIMA’s performance is competitive for one-day-ahead price forecasts where linear dynamics dominate \cite{xiaPredictionFinancialTimeseries2025,kareemBitcoinForecastingClassical2025,tangSurveyMachineLearning2022}, but accuracy degrades beyond that horizon.
\noindent \textbf{Generalised Autoregressive Conditional heteroskedasticity (GARCH):} Bollerslev's (1986) GARCH(1,1) models conditional variance as a function of past squared residuals and past conditional variance \cite{bollerslevGeneralizedAutoregressiveConditional1986}. Nelson (1991) identified standard GARCH treating positive and negative shocks symmetrically as a misspecification for assets with leverage effects \cite{nelsonConditionalHeteroskedasticityAsset1991}. 
These leverage effects are documented in Bitcoin by Katsiampa (2017), confirmed empirically by Chan et al. (2023) and Kareem et al. (2025) \cite{katsiampaVolatilityEstimationBitcoin2017, chanBitcoinHalvingCycle2023, kareemBitcoinForecastingClassical2025}.
\noindent Ardia et al. (2019) demonstrate that even a more flexible Markov-Switching GARCH requires five distinct states to adequately model Bitcoin's variance process, suggesting the regime structure is too complex for any pre-specified parametric model \cite{ardiaMarkovSwitchingGARCHModels2019}.

\noindent \textbf{Machine learning methods:} In their survey of ML models in FTS, Tang et al. (2022) find Logistic Regression (LR) to be used as a baseline in financial classification, particularly valued for its interpretability and well-calibrated probabilities. It is widely used in LR binary classification problems \cite{kalinicmilicevicWhatFactorsInfluence2023}. Compared to NNs, this model is less susceptible to overfitting thanks to its small number of parameters to estimate \cite{akyildirimPredictionCryptocurrencyReturns2021}.\\
SVM identifies an optimal separating hyperplane in a kernel-transformed feature space, making it well-suited to high-dimensional financial data with non-linear decision boundaries \cite{renForecastingStockMarket2019}. Empirically, in Omole and Enke’s study (2025), SVM outperformed LSTM and CNN-LSTM models (82\% F1-score vs. 66\% and 52\%) specifically on Bitcoin fragility classification using on-chain data, suggesting that feature quality and interpretable models can outperform architecturally complex networks when domain-relevant features are selected carefully \cite{omoleUsingMachineDeep2025}. This finding critically challenges the common assumption that deeper models necessarily generalise better on financial tasks. Despite requiring fewer computational resources than other prediction models, SVM needs more memory space to reserve all the generated support vectors \cite{tangSurveyMachineLearning2022}.\\
Random Forest (RF) is an ensemble of decision trees, each trained on a bootstrapped subsample of data, with predictions aggregated via majority vote \cite{breimanRandomForests2001}. In financial classification tasks, RF demonstrates resistance to overfitting through ensemble averaging and its ability to handle high-dimensional, non-linear feature spaces \cite{tangSurveyMachineLearning2022}. Jabbar \& Jalil (2024), benchmarking 41 models for algorithmic BTC trading, found that Random Forest and Stochastic Gradient Descent outperformed others in terms of profit and risk management \cite{jabbarComprehensiveAnalysisMachine2024}.\\
Gradient boosted trees with built-in L1/L2 regularisation ($reg_alpha$, $reg_lambda$) handle multicollinear feature sets well; sequential boosting corrects residual errors from previous trees; and they have built-in handling of missing values \cite{chenXGBoostScalableTree2016, hasanliIntegratingHighDimensionalTechnical2025, samsonDevelopmentBitcoinClosing2025}. Hasanli and Dursun (2025) found that XGBoost achieved the highest classification accuracy (55.9\%) for Bitcoin and the strongest trading performance (141.4\% cumulative return, Sharpe 1.78) across LR, RF, SVM, KNN, LSTM, and GRU \cite{hasanliIntegratingHighDimensionalTechnical2025}. Qureshi et al. (2025) verified that XGBoost had consistent predictive accuracy across varied market conditions, with F1-scores as high as 0.82 for some tests and ranked in the top 3 for random forest and gradient boosting for risk-adjusted metrics (1.30 Sharpe ratio and -9.0\% max drawdown) \cite{qureshiEvaluatingMachineLearning2025}.


\noindent \textbf{Deep learning methods:} Chen et al. (2024), in a state-of-the-art review of deep learning models for financial time series, confirm that hybrid CNN-LSTM and CNN-LSTM-Attention Mechanism (AM) architectures consistently outperform standalone models such as CNN-only or LSTM-only \cite{chenDeepLearningFinancial2024}. The LSTM architecture \cite{hochreiterLongShortTermMemory1997} addresses the vanishing gradient problem of standard RNNs through gating mechanisms that selectively retain long-term sequential dependencies—a critical advantage for Bitcoin price data, which exhibits multi-week bull/bear cycles. CNN layers preceding the LSTM function as a spatial feature extractor, capturing local temporal patterns (e.g., short-term momentum), while the LSTM handles longer-range dependencies. Chen et al. (2024) identify "domain knowledge negligence" (the tendency to treat financial prediction as a generic time-series problem without incorporating market structure knowledge) as a principal failing of deep learning approaches \cite{chenDeepLearningFinancial2024}.
% =============== DATA LABELLING METHODS ===============
\subsection{Data Labelling Methods}
Fixed-horizon binary labels (sign of N-day return) appeared to be the dominant labelling approach in literature. Comparative reviews by Jabbar \& Jalil (2024) and Qureshi et al. (2025) indicate that all papers reviewed by them used this scheme without explicit justification. The approach is simple, but ignores intra-period path, forces a position every day, and uses static thresholds that do not adapt to volatility \cite{qureshiEvaluatingMachineLearning2025, jabbarComprehensiveAnalysisMachine2024}.
Peng et al. (2024) suggest an auto-labelling approach to achieve reduction in transaction numbers while maintaining profitable performance. Their method first sequentially reduces the price series to a local minimum, smoothing out noise, then traverses the reduced series to assign three labels -- uptrend, downtrend, and stable -- based on fixed trend-change and stability-range thresholds \cite{pengAttentionbasedCNNLSTM2024}. However, the threshold values are static constants rather than volatility-scaled.\\
Regime-labelling via HMM offers a structurally different approach. Giudici \& Hashish and the AJPAS (2025) HMM paper fit 3-state HMMs (bull/neutral/bear) to Bitcoin price series using the Viterbi algorithm, then use the decoded hidden states as training labels. The appeal is that labels reflect persistent structural conditions rather than a single return outcome. The problem is that this is in-sample labelling – the HMM is fitted to the full price series, which introduces a severe look-ahead contamination problem \cite{machimboApplicationsHiddenMarkov2025}.\\
López de Prado (2018) addresses the path-dependency problem with Triple Barrier labelling: the label reflects which barrier is hit first (not just the endpoint), barriers are volatility-scaled, and the neutral class captures periods where no strong signal exists \cite{lopezdepradoAdvancesFinancialMachine2018}.
% =============== TRIPLE BARRIER LABELLING ===============
\subsubsection{Triple Barrier Labelling}
Triple barrier labelling (TBL), first introduced by de Prado \cite{lopezdepradoAdvancesFinancialMachine2018}, is a widely used method for generating labels in financial time series prediction by incorporating stop-loss, take-profit, and time horizons. Recent works have shown TBL's effectiveness in machine learning applications, achieving more balanced and realistic predictions compared to conventional labeling methods like raw returns and fixed time horizons \cite{liModelOptimizationStock2022}.\\
Using tick-level Bitcoin and Ethereum data from 2018 to 2023, Gradzki et al. (2025) compare triple barrier labelling with next-bar labelling under a walk-forward evaluation framework. Triple barrier produces better profitability and risk-adjusted returns (Sharpe ratios $\sim$2.0 compared to next-bar labelling, which showed higher accuracy but only 29 profitable experiments out of 210 \cite{gradzkiAlgorithmicCryptoTrading2025}.
Improvement of the Sharpe ratio is consistent across studies. Bieganowski and Slepaczuk (2024) establish that TBL-based setups improve the Sharpe and Information Ratio when combined with supervised autoencoders, evaluated on S\&P 500, EUR/USD, and BTC/USD data from 2010–2022, producing more actionable signals than raw prediction targets \cite{bieganowskiSupervisedAutoencoderMLP2024}.
Work by Fu et al.(2024) has shown TBL’s effectiveness in traditional machine learning applications. The study applies triple barrier labelling, optimised via an enhanced genetic algorithm. An XGBoost classifier trained on BTC/ETH November 2017–December 2023 pair trading data achieved a 51.42\% increase in profitability and a 73.24\% reduction in maximum drawdown compared with the traditional pair trading baseline \cite{fuEnhancedGeneticAlgorithmDrivenTriple2024}.
% =============== DATASET SPLITTING ===============
\subsection{Dataset Splitting}
The reality across Bitcoin machine learning projects is a lack of methodological consensus.\\
\textbf{Simple Chronological Holdout:} Experiments comparing 70/30, 80/20, and 90/10 splits by Ngai et al. (2023) for Support Vector Regression (SVR), Long Short Term Memory (LSTM), Extreme Gradient Boosting (XGBoost), and K-Nearest Neighbor Regression (KNN) models show lower RMSE as training data increases, indicating that model performance is sensitive to training-data size \cite{eddiengaiTimeSeriesPrediction2023}.A cryptocurrency classifier study by Qureshi et al. (2025) used 70/30 and 80/20 splits, calling the choice iterative \cite{qureshiEvaluatingMachineLearning2025}.\\
\textbf{Multiple folds Validation:} Cross-validation has been identified as a more robust process to improve generalisation \cite{dimitriadouPredictingBitcoinPrices2023}. Dimitriadou and Gregoriou used 5-fold repeated sampling within the training window specifically to stabilise hyperparameter estimates \cite{dimitriadouPredictingBitcoinPrices2023}.Jabbar and Jalil (2024) evaluate 41 ML models on Bitcoin, split into three distinct segments: training (Jan 2013 to Jan 2023), backtesting (Feb to Jul 2023), and forward testing (Aug to Oct 2023). Finally, the forward testing phase serves as a real-world application of the models; performance on unencountered data eliminates any survivorship bias. \\
\textbf{Systematically determined dataset split:} Yenidogan et al. (2018) use a systematic experimental approach to find the optimal data split for Bitcoin Forecasting using ARIMA and PROPHET for 2012-2018 data. A three-fold split (train, test, and validation) is used to avoid over-fitting. The optimal ratio for splitting data is established by observing the changes in the $R^2$, Mean Absolute Percentage Error (MAPE), and Root Mean Square Percentage Error (RMSPE) values as the training split increases at 10\% intervals. All values decreased to a minimum at a 70\% training set ratio before increasing, yielding optimal ratios of 70, 15, and 15 percent respectively \cite{yenidoganBitcoinForecastingUsing2018}.Oktem and Tekerek (2025) found that percentage-based dataset splits for training (56\%), validation (24\%), and testing (20\%) were optimised to prevent overfitting by flagging models with test RMSE $\geq 10\%$ validation RMSE  \cite{oktemBitcoinPriceDirection2025}
% =============== FEATURE SELECTION METHODS ===============
\subsection{Feature Selection Methods}
The majority of Bitcoin ML papers use no formal feature selection at all; features are chosen by domain intuition or prior literature and fed directly into the model \cite{chenAnalysisBitcoinPrice2023, raoComparativeStudyStatistical2025}. Filter methods rank features independently of any model using statistical measures. Wrapper methods evaluate feature subsets by training and scoring a model, making them model-aware at the cost of increased computational expense relative to filter methods.\\
Mutual Information method measures statistical dependence without assuming linearity. El Youssefi et al. (2025) systematically compared MI against RFE and Random Forest Importance across over 130 technical indicators for Bitcoin, Ethereum, and Binance Coin. MI selected momentum and volatility features most frequently; Bollinger Bands and True Range were consistently retained across methods. A key finding was that an 80–85\% reduction in the original feature set frequently maintains or enhances model performance \cite{youssefiOptimizingForecastAccuracy2025}. This is a noteworthy finding, showing that the majority of derived technical indicators convey similar information about the regime. This suggests a trade-off between adding dimensional complexity to data and extracting truly indicative signals. \\
Boruta is a frequently cited wrapper method in the Bitcoin-specific literature, it uses a Random Forest together with a feature importance measure to select the relevant feature from the input space \cite{kursaFeatureSelectionBoruta2010}. Dubey (2025) found that Boruta combined with CNN-LSTM performed better than L1 regression and Principal Component Analysis (PCA) methods for Bitcoin price direction prediction using on-chain data \cite{dubeyBitcoinPriceDirection2025}. Omole et al. (2025) applied Boruta to SVM, Random Forest, and deep learning models for Bitcoin price (range: 2013-2023) and movement prediction. Key finding was that Boruta feature selection method significantly improved the performance of the models, with Boruta-SVM emerging as the most profitable model (annual returns of 4249\%) \cite{omoleUsingMachineDeep2025}.\\
RFE removes the least important features in each iteration based on the model’s performance, providing a systematic way to identify the most relevant features \cite{youssefiOptimizingForecastAccuracy2025}. El Youssefi et al. (2025) found RFE primarily selected volatility and momentum indicators over the study's 1-20 day window, favoring short-term market shifts over trend and volume indicators \cite{youssefiOptimizingForecastAccuracy2025}.\\
Genetic Algorithm (GA) is a technique to generate and explore a search space using methods inspired by elements of nature \cite{contrerasGeneticAlgorithmFeature2024}. Contreras et al. (2024) applied GA-based feature selection to financial time series, using evolutionary search to find optimal feature subsets. This is the computationally most expensive method, typically only viable for smaller feature sets.\\
Feature Selection with Annealing (FSA) uses simulated annealing to explore the feature subset space more efficiently than exhaustive wrapper search. Pabuccu \& Barbu (2023) proposed FSA as an improvement over Boruta for financial data, outperforming it on all 10 datasets for regression and on 8 out of 10 for classification. For classification, FSA simplified the feature space more than Boruta (35 vs 88 features), showing better MSE (21.42 vs 24.82) and accuracy (0.525 vs 0.467) \cite{pabuccuFeatureSelectionAnnealing2024}.\\
Boruta is a rigorously validated wrapper method in the Bitcoin ML literature, with multiple studies confirming significant performance gains from its use, and its shadow-feature statistical test provides formal confirmation of feature relevance rather than a ranked list.
% =============== HYPERPARAMETER TUNING ===============
\subsection{Hyperparameter Tuning}
In a review of 14 Bitcoin machine learning studies from between (2018 and 2023), Kervanci et al. (2024) find grid search, random search, and Bayesian Optimisation as the most popular approaches \cite{kervanciBitcoinPricePrediction2024}.\\
Grid search is an optimisation technique that helps determine the optimum values with an exhaustive search over a manually defined discrete parameter grid \cite{kervanciBitcoinPricePrediction2024}. According to Yu and Zhu (2020), a simpler search technique leads to one of the most accurate estimates, as each parameter combination is run independently of the others \cite{yuHyperParameterOptimizationReview2020}. The critical weakness is the curse of dimensionality -- a grid over 5 parameters with 5 values each requires 3,125 evaluations. For complex models like XGBoost with 10+ parameters, this search becomes computationally expensive. Also, preliminary knowledge of the domain is required to set appropriate limits.\\
Random search samples hyperparameter combinations randomly from defined distributions. Bergstra \& Bengio (2012) proved empirically and theoretically that random search outperforms grid search for the same computational budget when only a subset of parameters meaningfully affect performance \cite{bergstra2012random}. Each hyperparameter’s search budget in grid search is constrained, a limitation that random search doesn't have. It performs better for non-uniform distributions of parameters. While still a computationally intensive method, it is better suited to narrow down the search space before using guided algorithms to find final values \cite{yuHyperParameterOptimizationReview2020}.\\
Bayesian Optimisation builds a probabilistic surrogate model of the objective function and uses it to select promising next trial points. Yu and Zhu (2020)  establish that this approach outperforms random search and grid search in two aspects: preliminary knowledge of the distribution of hyper-parameters is not required, and the second is posterior probability, which is the core idea of BO. Optuna is the most modern Bayesian framework. Used by Jabbar and Jalil (2024) with 100 trials per model as their tuning method, with PNL\% as the objective metric \cite{jabbarComprehensiveAnalysisMachine2024}. Akiba et al. (2019) benchmark Optuna's Tree-Structured Parzen Estimator (TPE) +Covariance Matrix Adaptation Evolution Strategy (CMA-ES) against Gaussian Process(GP)-based methods and random search. GPyOpt performed better than TPE+CMA-ES in 34/56 cases in finding the best-attained loss value, while Optuna converged an order of magnitude faster per trial \cite{akibaOptunaNextgenerationHyperparameter2019}. Bafna et al. (2025) attribute the slow processing time to Gaussian-process (GP) surrogates: GP hyperparameter fitting scales as $O(N^3)$ \cite{bafnaTakingGPOut2025}.
% =============== HYPERPARAMETER TUNING ===============
\subsection{Model Evaluation}
\textbf{Statistical metrics:} Samson et al. point out that the majority of Bitcoin classification papers report accuracy as the primary metric, which highlights a well-documented problem: accuracy is uninformative for imbalanced datasets and rewards models that simply predict the majority class \cite{samsonDevelopmentBitcoinClosing2025}. Studies by Hajare et al. (2025), Mallqui and Fernandes (2019), and Qureshi et al. (2025) support this argument by incorporating richer evaluation criteria like precision, recall, F1-score, and MCC to provide a comprehensive overview of performance \cite{qureshiEvaluatingMachineLearning2025, hajareBitcoinPricePrediction2025, mallquiPredictingDirectionMaximum2019}. A key limitation of the F1-score is that it does not account for true negatives, which matters in three-class problems where the neutral class represents a meaningful signal \cite{chiccoAdvantagesMatthewsCorrelation2020}.\\
MCC is the most information-theoretically complete single metric for classification—it accounts for all four quadrants of the confusion matrix and is robust to class imbalance \cite{jurmanComparisonMCCCEN2012}. It was explicitly recommended over accuracy and F1 for imbalanced datasets by Chicco \& Jurman (2020) \cite{chiccoAdvantagesMatthewsCorrelation2020}. However, its adoption in Bitcoin ML papers remains rare.\\
\textbf{Financial metrics:} Jabar and Jalil (2024) point out that accuracy metrics alone are not enough in algorithmic trading; the effectiveness of classification and regression models is assessed using a set of standard metrics. Financial metrics provide unique insights into the model’s risk management and overall economic viability \cite{jabbarComprehensiveAnalysisMachine2024}. \\
Sharpe ratio Is a dominant risk-adjusted metric in the literature \cite{jabbarComprehensiveAnalysisMachine2024, dubeyBitcoinPriceDirection2025, sebastiaoForecastingTradingCryptocurrencies2021} . It measures risk-adjusted return: annualised excess return divided by annualised volatility \cite{sharpeSharpeRatio1994}. Sortino ratio is a Sharpe variant that penalises only downside volatility \cite{DownsideRiskProQuest}. Superior to Sharpe for asymmetric return distributions, which Bitcoin is a strong characteristic given its crash-and-rally regime \cite{sheppertGTScoreRobustObjective2026}.\\
In a survey conducted by Sebastião and Godinho (2021) of 23 studies between 2015 and 2020, almost half the studies presented compared model performance to buy-and-hold, which is crucial for determining whether model skill actually exceeds the simplest possible baseline. \cite{sebastiaoForecastingTradingCryptocurrencies2021}