% !TEX root =  ../Report.tex

\section{Conclusion}
\label{sec:Conclusion}
This project investigated whether technical and on-chain features carry sufficient signal to predict Bitcoin price direction using a triple-barrier labelling framework. SVC with Max EV labels achieved a test Sortino ratio of 3.118 against a buy-and-hold benchmark of 2.542, whilst near-chance classification metrics across all models suggest the signal is narrow, regime-dependent, and inconsistent with strong-form predictability. The findings support a weak but exploitable departure from market efficiency under the specific 2023--2026 test regime, with the spread of labels identified as the primary driver of pipeline behaviour.

\subsection{Contributions}
This project makes three contributions to the application of financial machine learning to Bitcoin market fragility prediction.\\
First, a two-stage barrier optimisation procedure is proposed in which Optuna TPE searches over triple-barrier parameters constrained by a Shannon entropy floor, resolving the circular dependency between label construction and model learnability without requiring domain knowledge. The ablation comparison across two label configurations then becomes the central empirical contribution rather than an ancillary check.\\
Second, the project provides empirical evidence that barrier geometry propagates through the entire ML pipeline—altering Boruta feature rankings, Optuna-converged regularisation strengths, and confusion matrix failure modes—illustrated by the four-order-of-magnitude divergence in Logistic Regression's optimal regularisation parameter between configurations (C = 337.3 vs 0.027).\\
Third, a fully reproducible, notebook-structured pipeline is produced as an open artefact, integrating temporal leakage controls, MLflow experiment tracking, and trading performance evaluation against a passive benchmark across four classifiers and two label configurations on twelve years of Bitcoin data.

\subsection{Objectives Revisited}
All four Must objectives defined in Table \ref{tab:moscow} were met. Data acquisition and preprocessing (O1) produced a dataset of 69 features for 2014–2026, with the Fear and Greed Index excluded. The triple-barrier labelling scheme (O2) was implemented with EWMA volatility scaling, with Optuna TPE returning optimised label distributions for 2 objectives: balanced classes and maximum expected value (EV). All four classifiers were trained with time-series-aware cross-validation and class imbalance mitigation (O3). Trading and classification metrics were computed in full (O4). Both Should objectives were also met: Boruta feature selection (O5) confirmed a compact feature set without prior domain knowledge, and Bayesian hyperparameter search (O6) was completed for all model-configuration pairs with results logged via MLflow. The two Could objectives were not met within the project timeline: extended on-chain cycle indicators (O7) were out of scope, and full Purged K-Fold cross-validation (O8) was not implemented, remaining the primary recommendation for further work. O9 was not pursued by design.

\subsection{Recommendations for Future Work}
Future work should prioritise Purged K-Fold cross-validation to eliminate residual label leakage. The triple barrier search objective should be improved to a signed EV incorporating the $-1$ class to align it with the trading evaluation. Along with that, barrier calibration via genetic algorithm would offer more global exploration of the parameter space than Bayesian search. Hyperparameter optimisation would benefit from more tweaking of the search space. At the architectural level, ensemble methods combining the four base classifiers have shown superior out-of-sample stability in financial time series \cite{tanrikuluEffectDataTypes2025}. Finally, a multi-horizon label comparison beyond the fixed 7-day window remains an open design question.