% !TEX root =  ../Report.tex

\section{Ethics}
\label{sec: Ethics}
Machine learning applications in cryptocurrency markets raise ethical concerns that extend beyond the predictive performance of the model. This project operates at the intersection of three established areas of ethical risk: responsible data sourcing, algorithmic transparency, and the societal implications of predictive financial modelling. While the project required no formal ethics approval, the absence of a formal approval process does not diminish the obligation to identify and address wider ethical implications. All data is publicly available and no personal information was collected or processed, as confirmed in the PFS ethics submission. The sections below examine each risk area in turn, with reference to the design decisions taken during development that reflect awareness of these concerns.
% =============== Data Licensing and Provenance ===============
\subsection{Data Licensing and Provenance}
All data used in this project was sourced from publicly available, non-personal datasets. However, the absence of personal data does not eliminate data ethics obligations. Chawla et al. argue that data provenance—the traceable origin, transformation, and use of data within AI systems—is foundational to the FATE principles (Fairness, Accountability, Transparency, and Ethics) that govern responsible AI deployment; without it, the legitimacy of model outputs cannot be verified or audited \cite{memarianFairnessAccountabilityTransparency2023}. This project addresses provenance explicitly: on-chain metrics were obtained from the CoinMetrics Community archive, released under a Creative Commons CC BY-NC 4.0 licence, which permits non-commercial academic use provided attribution is given \cite{coinMetrics}. As CC BY-NC prohibits any profit-making use \cite{byBringingTransparencyData2025}, this dissertation falls within the permitted scope, though any future commercial deployment of the pipeline would require re-sourcing data under a compatible licence. Historical price data was retrieved via yfinance, which carries equivalent non-commercial restrictions in Yahoo Finance's terms of service \cite{YahooDeveloperAPI, YahooTermsService}. The FAIR (Findable, Accessible, Interoperable, Reusable) data principles advocate that research data should be as open as possible but as closed as necessary; both sources satisfy these criteria for academic purposes whilst appropriately limiting commercial exploitation \cite{wilkinsonFAIRGuidingPrinciples2016}.
% =============== Algorithmic Transparency and Bias ===============
\subsection{Algorithmic Transparency and Bias}
Random Forest and XGBoost, two of the four classifiers used in this project, are canonical black-box architectures; their predictions are not directly interpretable by a human practitioner without supplementary analysis. The inability to explain why a signal was generated poses a governance risk. Goodell et al. (2023) demonstrate that feature-importance-based explainability using Shapley-derived methods improves both interpretability and predictive accuracy in Bitcoin price modelling, establishing that transparency mechanisms and performance are not in tension \cite{goodellExplainableArtificialIntelligence2023}. The Shapley plots of all models are in Appendix \ref{subsec: Shap}.\\
A deeper form of bias arises from the dataset itself: the model is trained exclusively on Bitcoin data from 2014–2026, a single asset across a specific regulatory and macroeconomic regime. This project does not claim generalisability beyond Bitcoin, and any extension to other assets or time periods would require revalidation — an obligation the SAFE framework for quantitative trading (Sustainability, Accuracy, Fairness, Explainability) identifies as a minimum ethical standard for responsible ML deployment in financial markets \cite{SustainabilityAccuracyFairness}.
% =============== Data Licensing and Provenance ===============
\subsection{Responsible Deployment and Societal Impact}
Presenting the project’s results without contextualizing their limitations would constitute an irresponsible use of predictive modeling. The Association for Financial Markets in Europe (AFME) notes that AI trading algorithms, even when initially designed with controls against market abuse, can produce manipulative behavior through unintended emergent optimization \cite{AFMEConsiderationsEthical}. This project produces research outputs only, but acknowledging these risks reflects the professional obligation to consider downstream use.\\
Beyond model performance, this project is part of Bitcoin's broader environmental footprint. Stoll et al. (2019) calculated Bitcoin's annual electricity consumption at 45.8 TWh, with annual $CO_2$ emissions of 22.0–22.9 $MtCO_2$. This is equivalent to the emissions of Jordan or Sri Lanka \cite{stollCarbonFootprintBitcoin2019}. Fayza et al. (2025) demonstrate that CMOS-based computing no longer scales in energy efficiency according to Moore's Law, meaning the assumption that hardware improvements will neutralise growing ML and Bitcoin compute demands is no longer empirically supported \cite{fayzaPhotonicsSustainableAI2025}. Considering all the factors above, the decision to train on local hardware rather than cloud GPU was not only a budget constraint but also an ethical choice consistent with responsible engineering practice.
