% !TEX root =  ../Report.tex
\section{Discussion \& Analysis}
\label{sec: Discussion}

% =============== Triple Barrier Parameter Trade-off ===============
\subsection{Triple Barrier Parameter Trade-off}
The contour plot analysis (Figures \ref{fig:res tp contour} and \ref{fig:res tp contour entropy}) reveals a paradox: the regions of maximum Expected Value (profitability) and maximum Shannon Entropy (class balance) are geographically distinct in the parameter search space. High-entropy configurations (tighter barriers) provide the most balanced training signals, which will likely improve the F1 Macro score by reducing majority-class bias. However, these configurations coincide with a lower Expected Value, suggesting that the model is being trained to identify balanced but low-profitability signals. Conversely, the high-EV zone ($profit\_mult \approx 1.5$) exhibits lower entropy, indicating a drift toward class imbalance where the +1 label dominates.
The dense concentration of TPE trial points (black dots) in the upper-left quadrant of both contours confirms that the sampler correctly identified this constrained region as the feasible optimum given the entropy floor, rather than converging on the unconstrained global maximum that might produce degenerate label distribution.\\
The entropy floor at $H \geq 0.85$ is not a compromise imposed on the EV objective; it is the constraint that reduces the risk of failed F1 calculation due to division by zero when the minority class is missing from cross-validation (CV) folds, making the final model sensitive to the specific folds of data \cite{lopezdepradoAdvancesFinancialMachine2018}.\\
The practical consequence is that the label asymmetry imposed at this step propagates through the entire pipeline: Boruta's importance scores are conditioned on the label distribution they observe, which alters the selection of features that carry statistically significant predictive signals.
% =============== Feature Selection ===============
\subsection{Feature Selection}
The rejection of DailyReturn and LogReturn in Boruta selection results under both label configurations is consistent with the weak-form EMH at the daily frequency \cite{yiMarketEfficiencyCryptocurrency2023}. This finding aligns with Omole et al. (2025), who found that Boruta-selected feature subsets systematically excluded short-lag return features while retaining on-chain and momentum composites \cite{omoleUsingMachineDeep2025}. The dominance of CapMVRVCur across both configurations is consistent with the broader on-chain literature: the MVRV ratio encodes the aggregate unrealised profit/loss of the network and has a non-random predictive structure, shown by its widespread use in literature \cite{dubeyBitcoinPriceDirection2025, 230304642v1ForecastingMovements, palazziNetworkFundamentalsMacroFinancial2026, kanzariContextadaptiveIntelligentAgents2025}. Its wide IQR across Boruta iterations reflects regime-dependence rather than noise – the signal is real but its magnitude varies with market cycle phase, which is precisely the kind of non-stationary, episodic predictability the Adaptive Market Hypothesis (AMH) predicts \cite{karasinskiPredictabilityHighFrequencyReturns2025}.\\
The sensitivity of the feature set to the label proportions points to a bias intrinsic to the Mean Decrease Impurity (MDI) metric of Random Forest under skewed class priors \cite{stroblBiasRandomForest2007}. The five features confirmed under balanced entropy but rejected under Max EV (TxCnt, TxTfrCnt, FlowOutExNtv, TRANGE, and Momentum\_7day) are not necessarily less predictive; their predictive signal might be concentrated in the minority -1 class and may be less measurable by MDI when the stop-loss class is too sparse to anchor their impurity contribution \cite{HowAreFeature_importances10:28:55+00:00, stroblBiasRandomForest2007, janitzaAUCbasedPermutationVariable2013}. MDI evidence supports Kim and Kang's (2024) reasoning for balanced distributions: they lead to more robust multiclass models \cite{StockPricePrediction}.\\
% =============== Hyperparameter Tuning ===============
\subsection{Hyperparameter Tuning}
The four-order-of-magnitude divergence in Logistic Regression's regularisation parameter $C$ across label configurations (Max EV: 337.3; Balanced: 0.027) is a direct consequence of the label distribution shift rather than model instability. Under Max EV, the dominant class (+1, 44.25\% of training) produces a simpler training surface that Optuna exploits by removing regularisation—TPE is free to drive $C$ toward near-zero regularisation without incurring a CV penalty. Under Balanced labels, the strong regularisation by $C=0.027$ is the correct response to a decision boundary in a denser region of feature space, preventing the boundary from collapsing. Both configurations converge to a Lasso-dominant Elastic Net mix (l1\_ratio: 0.824 and 0.806, respectively), suggesting the underlying feature space is sparse. \\
SVM-RBF selects a twofold higher C under balanced labels (76.71 vs. 30.63) alongside a lower gamma ($4.51\times10^{-4}$vs. $3.01\times10^{-3}$). The wider margin required under balanced labels is achieved by simultaneously softening the penalty on support vectors and broadening the kernel bandwidth, which is consistent with a more complex, less separable decision boundary. \\
Random Forest, with a near-maximal min\_samples\_leaf under both configurations (19 and 18), acts as the primary regulariser against the noisy financial labels. The shallow max\_depth of 3 selected under balanced labels and trees with max\_depth 11 are not straightforwardly interpretable as an optimal structural choice; they contradict the patterns shown by other models. For the maximum Expected Value (EV) configuration, a short profit-margin multiplier means that, due to the overlapping distinction between class +1 and class 0 in the feature space, TPE explores deeper trees (depth 11) to find these fine-grained boundaries. The objective landscape may be flat for the balanced configuration, where multiple depth values yield equivalent F1-macro scores \cite{watanabeTreeStructuredParzenEstimator2023}.\\
XGBoost exhibits opposing regularisation regimes across configurations, consistent with Logistic Regression results. The (learning\_rate, n\_estimators) pairs are internally consistent across both configurations. Max EV selects fast convergence with few trees (0.074, 100) and Balanced selects slow convergence with many trees (0.010, 450), confirming that TPE correctly explored this classically coupled interaction. Max EV is a simpler problem with an effective two-class boundary (+1 vs. 0), a lower-complexity problem.
% =============== Model Classification Performance ===============
\subsection{Model Classification Performance}
The contrasting generalisation profiles of Random Forest and XGBoost under Max EV labels illustrate two distinct failure modes. Random Forest's deep trees (max\_depth 11) with heavy leaf constraints (min\_samples\_leaf 19) memorised a rich three-class training surface -- train MCC 0.710 -- that collapsed entirely on the test set (MCC 0.055), yielding the largest overfit gap of any model (0.618). XGBoost avoided this memorisation by converging to a simpler two-class boundary (100 trees, fast learning rate, strong L2), producing a smaller overfit gap -- but at the cost of systematically ignoring the stop-loss class, yielding near-zero test MCC (0.002). These represent opposite ends of the bias-variance tradeoff: Random Forest memorised the noise in the training surface while XGBoost settled for an incomplete one.\\
SVC achieves the best test MCC (0.070) precisely because its maximum-margin objective neither memorises training noise nor discards minority-class structure. A finding consistent with Omole et al. (2025), who attributed SVM’s success to the RBF kernel's margin-maximising geometry’s resistance to label noise \cite{omoleUsingMachineDeep2025}.

